Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-03-28 GWLDCI.tex
intro to first extended

In recent years, there has been a an active program researching the geometric and combinatorics underlying SYM $N=4$ theory \cite{wilsonloops, amplituhedron, correlahedron, karpwilliams, galashinlam}. This body of work started with the observation that the on shell amplitudes of this theory correspond to the volume of a subspace of a positive Grasmannian, called an Amplituhedron \cite{amplituhedron}. Since then, there has been significant work to study the structure of the amplituhedron, geometrically and combinatorially, for instance \cite{amplituhedron2, karpwilliams, galashinlam}.

Meanwhile, there has been a different body of work studying SYM $N=4$ theory from the point of view of Feynman integrals \cite{BullimoreSkinner, AldayMaldecena, others}. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and the positive Grassmannians \cite{wilsonloops}. Others have tried to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{correlahedron, non-orientable}.

In this paper, we expand on map between Wilson loop diagrams and positroids, first defined in \cite{wilsonloops}. Each Wilson loop diagram, $W$, has associated to it a real-variable valued matrix $C(W)$, which defines positroid, i.e. a matroid that can be realized as an element of $\Gr(k,n)$. The subspace of $\Gr(k,n)$ parametrized by each $C(W)$ is the positroid cell defined by $W$, called $\Sigma(W)$. Lemma ?? of \cite{wilsonloops} identifies that certain Wilson loop diagrams define the same positroid cells. In particular, the authors define a subset of Wilson loop diagrams, called exact diagrams, and show that if two Wilson loop diagrams only differ by an exact subdiagram, they define the same positroid cell.

In the current work we study exact subdiagrams in greater detail, identifying several matroidal properties. Furthermore, we enumerate how many different Wilson loop diagrams map to any positroid cell, $\Sigma(W)$. Furthermore, we show that counting the number of distinct $\Sigma(W)$, positroid cells that correspond to Wilson loop diagrams, corresponds to counting the number of directions of an associahedron.

>>>>

In recent years, there has been a an active program researching the geometric and combinatorics underlying SYM $N=4$ theory \cite{wilsonloops, amplituhedron, correlahedron, karpwilliams, galashinlam}. This body of work started with the observation that the on shell amplitudes of this theory correspond to the volume of a subspace of a positive Grasmannian, called an Amplituhedron \cite{amplituhedron}. Since then, there has been significant work to study the structure of the amplituhedron, geometrically and combinatorially, for instance \cite{amplituhedron2, bigamplituhedronpaper, karpwilliams, galashinlam}. The focus of much of this work has been to understand the relationship between the combinatorial objects from the physics of the system, BCFW diagrams, and their corresponding natural object for studying the positive Grassmannians, positroid cells.

Meanwhile, there has been a different body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space\cite{BullimoreSkinner, AldayMaldecena?, Wittenpaper}. These integrasl are calculated via Wilson loops. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloops}. With the second author and others, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmanians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{correlahedron, HeslopStewart}.

Both the Amplituhedron literature and the Wilson loop literature associate $n$ particle, $N^kMHV$ interaction in SYM $N=4$ theory to volume forms on a set of positroid cells of $\Gr(k,n)$. This gives a CW complex of some submanifold in $\Gr(k,n)$. Understanding the geometry and topology of this space is the subject of ongoing research. For Wilson loop diagrams with $k=2$ and $n=6$, one knows that the manifold is not contractible \cite{casestudy}. Furthermore, the spaces defined by the Amplituhedron and the Wilson loop diagrams are not the same \cite{correlahedron}. In particular, it it more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams. In this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}.

In this paper, we focus on the correspondence between Wilson loop diagrams and positroids, first defined in \cite{wilsonloops}. In that paper, the authors show that there are certain families of diagrams, specifically those with exact subdiagrams, that do map to the same positroid cell. In \cite{casestudy} the authors explictly show that not every positroid cell of a given dimension corresponds to a Wilson loop diagram. In this paper, we better enumerate this correspondence. We show that the map from Wilson loop diagrams to postroid cells is one to one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We show that exact diagrams correspond to uniform matroids, and that contraction by exact subdiagrams has special matroidal properties. Further, given a Wilson loop diagram with a non-trivial exact subdiagram, we count the total number of Wilson loop diagrams that correspond to that same positroid cell. While we do not count the number of distint positroid cells that correspond to Wilson loop diagrams, we do show that this enumeration is equivalent to enumerating the number of directions of an associahedron.
